### 2024-09-25 paper.tex
abstract

\documentclass[conference]{resources/IEEEtran}

%%% Start Preamble
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\AtBeginDocument{%
	\providecommand\BibTeX{{%
			\normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\usepackage{hyperref}
\usepackage{cite}

\usepackage{amsmath,amssymb,amsfonts} %amsmath already included
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx, algorithm}
\usepackage{graphicx} %graphicx already included
\usepackage{textcomp}
\usepackage{xcolor}

\usepackage{tikz}
\usetikzlibrary{arrows,decorations.markings, positioning}
\usepackage{subcaption}
\usepackage{tabu}

\hyphenation{con-straints}
%\pagestyle{plain}

%% COMMANDS:
\usepackage{resources/mathcommands}

%% DRAW SCHEDULES:
\usepackage{resources/scheduling}

%% THEOREMS:
% Theorems --
\usepackage{amsthm}

\newtheorem{thm}{Theorem}%[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{obs}[thm]{Observation}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{notn}[thm]{Notation}

\theoremstyle{remark}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{exmpl}[thm]{Example}
% --

%% WRITING:
\newcommand{\comment}[3]{{\color{#1}#2:#3}\ClassWarning{}{There are comments left in the paper. [Remove them before submission!]}}
\newcommand{\mario}[1]{\comment{red}{mg}{#1}}
\newcommand{\kuan}[1]{\comment{blue}{kh}{#1}}
\newcommand{\jj}[1]{\comment{orange}{jj}{#1}}

\usepackage{lineno}
%\usepackage[switch]{lineno}
\linenumbers
% fix problem with line numbering around align
\makeatletter
\let\LN@align\align
\let\LN@endalign\endalign
\renewcommand{\align}{\linenomath\LN@align}
\renewcommand{\endalign}{\LN@endalign\endlinenomath}
\makeatother

%% PAPER SPECIFIC:

%%% End Preamble

\begin{document}
	%% TITLE:
	\title{Optimal Task Phasing for End-To-End Latency in Semi-Harmonic Task Systems}

	%% AUTHORS: Mario, Matthias

%	% Final version
%	\author{
%		\IEEEauthorblockN{First Author, Second Author and Third Author}
%		\IEEEauthorblockA{TU Dortmund University\\
%			Email: \{first.author, second.author, third.author\}@tu-dortmund.de}
%		\and
%		\IEEEauthorblockN{Other Author}
%		\IEEEauthorblockA{Other University\\
%			Email: other.author@other-university.de}
%		}

%	% Submission version:
	\author{RTAS 2025 Submission \# \textcolor{red}{TBD} \quad\quad\quad Pages: \pageref{last-page}}

	\maketitle

	\begin{abstract}
		% - introductory sentence to the topic (optional)
		In the context of automotive systems, the end-to-end latency of a sequence of tasks (a so-called cause-effect chains) is a common metric to ensure correct timing behavior.
		% - describe the issue or main problem (create a need)
		To control the end-to-end latency, proper task configuration is crucial. 
		While the literature considers the configuration of task periods, optimization of task phases to minimize the end-to-end latency is only sparsely discussed.

		% - one sentence detailing your approach (meet the need)
		In this work, we discuss the configuration of task phases to optimize the end-to-end latency of a cause-effect chain.
		% - list main research results
		To that end, we develop a strategy for semi-harmonic task systems, which are very common in industrial applications.
		% - finish with MAIN MESSAGE of the paper
		We prove that our strategy is optimal in the sense that it minimizes the end-to-end latency.
		Furthermore, our evaluation based on a real-world application shows that optimizing task phases can reduce end-to-end latencies significantly. 
	\end{abstract}

%	\vspace{-.2in}

	\section{Introduction}
	\label{sec:introduction}
	% General introduction

	% Topic of this paper

	% Detailed topic / tell more details

	% Contribution
	\noindent\textbf{Contributions:} % short text
	\begin{itemize}
		\item one
		\item two
		\item three
	\end{itemize}

\section{System Model}
\label{sec:system_model}

\section{Problem Definition}
\label{sec:problem_def}

\section{Main Topic}

\section{Evaluation}
\label{sec:evaluation}

%\vspace{-0.05in}
\section{Conclusion}
\label{sec:conclusion}

% bib
\label{last-page}
\clearpage
%\bibliographystyle{abbrv}
%\bibliography{real-time}

\end{document}